---
title: "Homework 10"
author: "Maria Alfaro"
date: "3/30/2022"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

[Home Page](index.html)

#### 1. Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.

```{r}
counter = 0
pre_vec = runif(0:100)
my_vec = ifelse(pre_vec>0.5,1,0)
print(my_vec)

count_myvec = function(counter=0){
  for(i in 1:length(my_vec)){
  if(my_vec[i]==0){
    counter = counter + 1
  }
  }
  return(counter)
}
count_myvec()
```

#### 2. Use subsetting instead of a loop to rewrite the function as a single line of code.
```{r}
count_myvec2 = function(counter=0){
  length(my_vec[my_vec == 0])
}
count_myvec2()
```

#### 3. Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.
```{r}
timestable = function(x=9,y=10){
  makeTT = matrix(nrow=y,ncol=x)
  for(i in 1:nrow(makeTT)){
    for(j in 1:ncol(makeTT)){
      makeTT[i,j] = i*j
    }
  }
return(makeTT)  
}
ninebyten = timestable()
print(ninebyten)
```

#### 4. In the next few lectures, you will learn how to do a randomization test on your data. We will complete some of the steps today to practice calling custom functions within a for loop. Use the code from the March 31st lecture (Randomization Tests) to complete the following steps:

##### a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.

```{r}
group1 = rnorm(n=20,mean=5,sd=0.1)
group2 = rnorm(n=20,mean=10,sd=0.1)
group3 = rnorm(n=20,mean=15,sd=0.1)

group = c(rep("group1",20),rep("group2",20),rep("group3",20))
response = c(group1,group2,group3)

my_df = data.frame(group,response)

print(my_df)

```

##### b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.

```{r}
library(dplyr)

shuffle_calcmeans = function(df=my_df){
  shuff_resp = sample(df$response)
  shuff_df = data.frame(group,shuff_resp)
  group_means = aggregate(shuff_df$shuff_resp,list(shuff_df$group),FUN=mean)
  vec_means = group_means$x
  
  return(vec_means)
}

shuff_means = shuffle_calcmeans()   
print(shuff_means)
  
```

##### c. Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.
```{r}
hun_shuff_df = data.frame(replicate=NA,group1_mean=NA,group2_mean=NA,group3_mean=NA)
  
for(i in 1:100){
  hun_shuff_df[i,] = c(i,shuffle_calcmeans())
}

```


##### d. Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?

```{r}
library(ggplot2)
library(tidyr)

final_df = pivot_longer(hun_shuff_df,c("group1_mean","group2_mean","group3_mean"),names_to="group",values_to="means")
final_df = select(final_df,-"replicate")

hun_shuff_plot = ggplot(final_df,aes(x=means,fill=group)) +
  theme_bw(12) +
  geom_histogram(color='#e9ecef',alpha=0.4,position='identity',bins=30)
  

hun_shuff_plot

```
The reshuffled means are all around 10, whereas the original means were more distinctly different (5,10,15).
