---
title: "Homework 4"
author: "Maria Alfaro"
date: "2/9/2022"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

[Home Page](index.html)

## Problem 1
**Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.**

x^a^b   
(xa)^b   
3x^3+2x^2+1

```{r}
x = 1
a = 2.2
b = 3.3
```

```{r}
z = x^a^b
print(z)

```

```{r}
z = (x^a)^b
print(z)

```

```{r}
z = 3*(x^3) + 2*(x^2) + 1
print(z)

```

## Problem 2
**Using the rep and seq functions, create the following vectors:

(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)   
(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)   
(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)**

```{r}
s = c((seq(from=1,to=8)), (seq(from=7,to=1)))
print(s)
```

```{r}
s = c(rep(1,1),rep(2,2),rep(3,3),rep(4,4),rep(5,5))
print(s)

```

```{r}
s = c(rep(5,1),rep(4,2),rep(3,3),rep(2,4),rep(1,5))
print(s)

```

## Problem 3
**Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates.**

```{r}
n = c(runif(2))
print(n)

y = atan(n[2]/n[1])
print(y)
x = sqrt((n[1]^2)+(n[2]^2))
print(x)

```
**x=r, y=theta**

## Problem 4
**Create a vector queue <- c("sheep", "fox", "owl", "ant") where queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update queue as:**

the serpent arrives and gets in line;   
the sheep enters the ark;   
the donkey arrives and talks his way to the front of the line;   
the serpent gets impatient and leaves;   
the owl gets bored and leaves;   
the aphid arrives and the ant invites him to cut in line.   
Finally, determine the position of the aphid in the line.

```{r}
queue = c("sheep", "fox", "owl", "ant")
print(queue)

```

```{r}
queue = append(queue, "serpent")
print(queue)

```

```{r}
queue = queue[-1]
print(queue)

```

```{r}
queue = append(queue, "donkey", after = 0)
print(queue)

```

```{r}
queue = queue[-5]
print(queue)

```

```{r}
queue = queue[-3]
print(queue)

```

```{r}
queue = append(queue, "aphid", after = 2)
print(queue)

```
**The aphid is 3rd in line!**

## Problem 5
**Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.**

```{r}
v = c(seq(from=1,to=100,by=1))

v2 = which(v %% 3 != 0 & v %% 7 != 0 & v %%  2 != 0)

print(v2)
```





