---
title: "Homework 6"
author: "Maria Alfaro"
date: "2/23/2022"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---

```{r}
library(ggplot2)
library(MASS)
library(tidyverse)

z = read_csv("LTER_RandomLake.csv")
str(z)
summary(z)

```

## Plot histogram of data
```{r}
p1 = ggplot(data=z[which(z$pH>-900),], aes(x=pH, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```
## Add empirical density curve
```{r}
p1 =  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```
## Get maximum likelihood parameters for normal
```{r}
normPars = fitdistr(z$pH,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"]
```
## Plot normal probability density
```{r}
meanML = normPars$estimate["mean"]
sdML = normPars$estimate["sd"]

xval = seq(0,max(z$pH),len=length(z$pH))

stat = stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$pH), args = list(mean = meanML, sd = sdML))
p1 + stat
```
## Plot exponential probability density
```{r}
expoPars = fitdistr(z$pH,"exponential")
rateML = expoPars$estimate["rate"]

stat2 = stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$pH), args = list(rate=rateML))
p1 + stat + stat2
```
## Plot uniform probability density
```{r}
stat3 = stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$pH), args = list(min=min(z$pH), max=max(z$pH)))
p1 + stat + stat2 + stat3
```
## Plot gamma probability density
```{r}
gammaPars = fitdistr(z$pH,"gamma")
shapeML = gammaPars$estimate["shape"]
rateML = gammaPars$estimate["rate"]

stat4 = stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$pH), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```
## Plot beta probability
```{r}
pSpecial = ggplot(data=z, aes(x=pH/(max(pH + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars = fitdistr(x=z$pH/max(z$pH + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML = betaPars$estimate["shape1"]
shape2ML = betaPars$estimate["shape2"]

statSpecial = stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$pH), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```

## Find best-fitting distribution
```{r}
meanML = normPars$estimate["mean"]
sdML = normPars$estimate["sd"]

normal = data.frame(normie=rnorm(181,mean=meanML,sd=sdML))
normal

normplot = ggplot(data=normal, aes(x=normie, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(normplot)

normstat = stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(normal$normie), args = list(mean = meanML, sd = sdML))
normplot + normstat

p1 + stat

```

## How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?

It actually does a pretty good job! pH must be pretty "normal" ;)

